---
title: "Getting Started with ParametricSurvFit"
author: "Gosuke Homma"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with ParametricSurvFit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Set to TRUE when you have actual data
)
```

## Introduction

The `ParametricSurvFit` package provides a comprehensive suite of functions for fitting parametric survival distributions to clinical trial data. This vignette demonstrates the main functionality using ADaM datasets commonly used in pharmaceutical research.

The package now includes visualization capabilities with Kaplan-Meier curves and parametric distribution overlays, making it a complete solution for parametric survival analysis.

## Installation

```{r installation}
# Install from GitHub
devtools::install_github("gosukehommaEX/ParametricSurvFit")

# Load the package
library(ParametricSurvFit)
```

## Overview of Functions

The package provides five main functions:

1. **`DataParametricSurv()`**: Prepares ADaM datasets for survival analysis
2. **`FitSurvMods()`**: Fits multiple parametric distributions and creates formatted tables
3. **`ExtractParams()`**: Extracts parameters from individual fitted models
4. **`ExtractCovarianceMatrix()`**: Extracts variance-covariance matrices from fitted models
5. **`ParametricSurvKM()`**: 🆕 Creates Kaplan-Meier curves with parametric overlays

## Basic Workflow

### Step 1: Prepare the Data

```{r data_prep}
# Basic usage - Overall Survival analysis with sex stratification
surv_data <- DataParametricSurv(
  adsl_path = "path/to/adsl.sas7bdat",
  adtte_path = "path/to/adtte.sas7bdat",
  population = "ITTFL",      # ITT population
  variable = "OS",           # Overall Survival
  stratify_by = "SEX"        # Stratify by sex
)

# View the structure of the prepared data
str(surv_data)
head(surv_data)
```

### Step 2: Fit Parametric Models

```{r fit_models}
# Fit all available distributions
results_all <- FitSurvMods(
  dataset = surv_data,
  table_caption = "Parametric Models for Overall Survival by Sex"
)

# Display the formatted table
results_all
```

### Step 3: Extract Covariance Matrices

```{r covariance_analysis}
# Extract covariance matrix for Weibull distribution
cov_weibull <- ExtractCovarianceMatrix(
  dataset = surv_data,
  distribution = "weibull",
  parameter_name = "Overall Survival",
  stratify_reference_level = "M",  # Use Male as reference
  create_kable = TRUE
)

# Display the formatted covariance matrix
print(cov_weibull)
```

### Step 4: Create Kaplan-Meier Visualizations 🆕

```{r km_visualization}
# Create KM plots with Weibull overlay
km_plots <- ParametricSurvKM(
  dataset = surv_data,
  distribution = "weibull",
  figure_caption = "Overall Survival Analysis",
  stratify_name = "SEX",
  control_arm = "Placebo",
  time_scale = "Months"
)

# Display plots (will show separate panels for Male and Female)
print(km_plots)

# Access individual plots
male_plot <- km_plots[["SEX_M"]]
female_plot <- km_plots[["SEX_F"]]

# Save individual plots
# ggsave("OS_Male_weibull.png", male_plot$plot, width = 12, height = 8)
# ggsave("OS_Female_weibull.png", female_plot$plot, width = 12, height = 8)
```

## Advanced Examples

### Example 1: Complete Parametric Analysis Pipeline

```{r complete_pipeline}
# 1. Prepare data for PFS analysis by region
pfs_data <- DataParametricSurv(
  adsl_path = "path/to/adsl.sas7bdat",
  adtte_path = "path/to/adtte.sas7bdat",
  population = "SAFFL",
  variable = "PFS",
  stratify_by = "REGION"
)

# 2. Fit multiple distributions
pfs_results <- FitSurvMods(
  dataset = pfs_data,
  distributions = c("exp", "weibull", "gamma"),
  table_caption = "PFS Analysis by Region"
)

# 3. Extract covariance for best model
cov_gamma <- ExtractCovarianceMatrix(
  dataset = pfs_data,
  distribution = "gamma",
  include_shape = TRUE,
  stratify_reference_level = "US",
  parameter_name = "Progression-Free Survival"
)

# 4. Create visualizations with custom settings
pfs_plots <- ParametricSurvKM(
  dataset = pfs_data,
  distribution = "gamma",
  figure_caption = "Progression-Free Survival by Region",
  stratify_name = "REGION",
  control_arm = "Control",
  time_scale = "Weeks",
  time_max = 100
)

# Display results
print(pfs_results)
print(cov_gamma)
print(pfs_plots)
```

### Example 2: Covariance Matrix with Shape Parameters

```{r advanced_covariance}
# Gamma with shape parameter included
cov_gamma_shape <- ExtractCovarianceMatrix(
  dataset = surv_data,
  distribution = "gamma",
  include_shape = TRUE,
  stratify_reference_level = "F",  # Use Female as reference
  parameter_name = "Overall Survival with Shape",
  create_kable = TRUE
)

print(cov_gamma_shape)
```

### Example 3: Comparing Multiple Distributions Visually

```{r multiple_distributions_visual}
# Compare different distributions visually
distributions_to_compare <- c("exp", "weibull", "gamma")

# Create plots for each distribution
comparison_plots <- list()

for (dist in distributions_to_compare) {
  comparison_plots[[dist]] <- ParametricSurvKM(
    dataset = surv_data,
    distribution = dist,
    figure_caption = paste("Overall Survival -", toupper(dist), "Distribution"),
    stratify_name = "SEX",
    control_arm = "Placebo",
    time_scale = "Months",
    return_plots = TRUE
  )
}

# Display comparison
for (dist in names(comparison_plots)) {
  cat("\n=== Plots for", toupper(dist), "Distribution ===\n")
  print(comparison_plots[[dist]])
}
```

### Example 4: Customizing Visualization Parameters

```{r custom_visualization}
# Create highly customized plots
custom_plots <- ParametricSurvKM(
  dataset = surv_data,
  distribution = "weibull",
  figure_caption = "Overall Survival Analysis - Custom Settings",
  stratify_name = "SEX",
  control_arm = "Control",  # Specify control arm
  time_scale = "Years",     # Use years instead of months
  time_max = 5,            # Limit to 5 years
  return_plots = TRUE      # Return plot objects
)

# Customize further if needed
# custom_plots[["SEX_M"]]$plot <- custom_plots[["SEX_M"]]$plot + 
#   ggplot2::theme(legend.position = "bottom")
```

## Key Features of ParametricSurvKM

The new `ParametricSurvKM()` function provides:

### Visualization Features
- **Professional KM curves** using the survminer package
- **Parametric overlays** as dashed lines matching treatment arm colors
- **Risk tables** properly aligned with the survival curves
- **Parameter annotations** showing distribution estimates on each plot

### Customization Options
- **Color coding**: Control arm (red), Treatment arm (blue)
- **Time scales**: Flexible axis labeling (Months, Years, Weeks, Days)
- **Stratified panels**: Separate plots for each stratification level
- **Custom captions**: Personalized figure titles and subtitles

### Export Capabilities
- **ggplot2 objects**: Publication-ready plots
- **Individual access**: Extract specific stratified plots
- **High-quality output**: Suitable for presentations and publications

## Tips for Best Results

1. **Choose appropriate distributions**: Start with exponential and Weibull for most analyses
2. **Use meaningful stratification**: Sex, region, and age groups are common choices
3. **Set proper reference levels**: Use the most common or clinically relevant category
4. **Customize time scales**: Match your study's reporting conventions
5. **Save plots appropriately**: Use adequate resolution and size for your intended use

## Session Information

```{r session_info}
sessionInfo()
```
